Testing Charts: viewer's perceptual accuracy in surveys

### 2023-07-10 SDSS-2023/testing-charts-accuracy.qmd
(1) area of wedges vs tiles (2) effective number of responses for rounds 6 and 7
changed template for JDS ... not quite there yet but getting closer

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE, library_id=rbbt::bbt_library_id('Graphics Research'), translator='bibtex')`"
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
format: 
  pdf:
    include-in-header:
      - edits.tex
      - authors.tex
      # - refs.tex
    keep-tex: true
    # template: template.tex
documentclass: jds
---

@@

\hh{This is Heike's color for making changes}
## Introduction

@@

|                                   |                                   |
|-----------------------------------|-----------------------------------|
| **Unaligned positions**           | **Aligned positions**             |
| ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) |

: The two stacked barcharts every participant got to see. In each barchart, the two marked tiles are to be compared for their size. in both instances, the tile on the right is (very slightly) larger. {#tbl-tasks-12 tbl-colwidths="\[50,50\]"}

@@

The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey. Rounds 1 and 2 are combined by adjusting the weights  with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`. Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis. 
Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).

>>>>

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: '1,2'               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Brett
    first-init: B.
    last: Berger
    marker: 2
    footnote-id: 2
  - first: Carl
    first-init: C.
    last: Camp
    marker: 3
  - first: David
    first-init: D.
    last: Dodge
    marker: 3
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
  - content: Another footnote.
    id: 2
affiliation:
  - institution: University of Achievement
    prefix: Department A        # optional
    country: Country A
    marker: 1
  - institution: Institution B
    country: Country B
    marker: 2
  - institution: Institution C
    country: Country C
    marker: 3
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
format:
  pdf:
    template: template-jdsart.cls.tex
    include-in-header: 
      - text: |
          \usepackage{amsfonts,amsmath,amssymb,amsthm}
          \usepackage{booktabs}
          \usepackage{graphicx}
          \usepackage{longtable}
          \usepackage{hyperref}
      - file: edits.tex
    cite-method: natbib
---

<!--
\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\setlength{\parindent}{0pt}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
  \usepackage{booktabs}
  \usepackage{lipsum}
  \usepackage{subfig}
  \usepackage{graphicx}
  \usepackage{longtable}
\hh{hello world!}
-->

```{r run-to-update-bibfile, eval=FALSE, include=FALSE}
keys <- rbbt::bbt_detect_citations("testing-charts-accuracy.Rmd")
citations_only <- grep(pattern="[fig|tbl|eq][-|:].*", keys, value=TRUE, invert=TRUE)
rbbt::bbt_write_bib(keys=citations_only, 'references.bib', overwrite = TRUE,
                    library_id=rbbt::bbt_library_id('Graphics Research'),
                    translator='bibtex')
```

@@

## Introduction

@@

```{r}
#| label: tbl-tasks-12
#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e. A and C are of identical size, as are B and D. While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
#| tbl-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout: [[45,45], [100]]
#| fig-align: "center"
#| out-width: 100%
knitr::include_graphics("../SDSS-2023/images/vertical-ab.PNG")
knitr::include_graphics("../SDSS-2023/images/vertical-cd.PNG")
```

<!-- |                                   |                                   | -->
<!-- |-----------------------------------|-----------------------------------| -->
<!-- | **Unaligned positions**           | **Aligned positions**             | -->
<!-- | ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) | -->

<!-- : The two stacked barcharts every participant got to see. In each barchart, the two marked tiles are to be compared for their size. in both instances, the tile on the right is (very slightly) larger. {#tbl-tasks-12 tbl-colwidths="\[50,50\]"} -->

@@

The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey. Rounds 1 and 2 are combined by adjusting the weights  with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`. Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis. 
Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).
